---
title: "R Notebook"
output: html_notebook
---
## Load libraries
```{r}
library(tidyverse)
library(maxLik)
library(survival)
```

## Load the data
```{r}
library(asaur)
dat <- pbc

# recode status&sex
dat$status = ifelse(dat$status == 2, 1, 0)
dat$sex = ifelse(dat$sex == "f", 1, 0)
table(dat$status)/length(dat$status)

# convert the time in days to years for simplicity
dat$time = floor(dat$time / 365)

# remove all rows with missing values from the dataframe
dat <- dat[complete.cases(dat), ]
head(dat)
```
## Histogram of age & sex distribution
```{r}
# Create a histogram of age
ggplot(dat, aes(x=age)) +
    geom_histogram(binwidth=2, colour="black", fill="pink") + ggtitle("Distribution of Age") + xlab("Age (years)") +  ylab("Frequency") +   ggtitle("Histogram of Age") + 
  theme(plot.title = element_text(hjust = 0.5)) + scale_x_continuous(breaks = seq(0, 80, by = 3))
summary(dat$age)
table(d$sex)

```
## Correlation Matrix
```{r}
cor.mtest <- function(mat) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j])
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}
# matrix of correlation statistic
corr<-cor(dat)
# matrix of the p-value of the correlation
p.mat <- cor.mtest(dat)

# Specialized the insignificant value according to the significant level
corrplot(corr, type="upper", order="hclust", 
         p.mat = p.mat, sig.level = 0.05)  

```
## Density Distribution of Quantitative Varibles
```{r}
xlabels <- list("bili" = "Serum Bilirubin (mg/dl)", 
                "chol" = "Serum Cholesterol in (mg/dl)", 
                "albumin" = "Albumin (gm/dl)", 
                "copper" = "Urine Copper (ug/day))", 
                "alk.phos" = "Alkaline Phosphatase (U/liter))", 
                "ast"=	"aspartate aminotransferase, once called SGOT (U/ml)", 
                "trig" = "triglicerides in mg/dl platelet = platelets per cubic ml/1000", 
                "protime" = "prothrombin time in seconds")
density_plot<-function(x, log = FALSE){
  if(log){
    return(ggplot(dat, aes(x=log(dat[[x]]))) + geom_density() + xlab(paste0("log(dat$",x,")")) +  ylab("Density") + ggtitle(paste0("log(dat$", x,")")) + theme(plot.title = element_text(hjust = 0.5)))
  }
  return(ggplot(dat, aes(x=dat[[x]])) + geom_density() + xlab(xlabels[[x]]) +  ylab("Density") + ggtitle(paste0("Density Plot of dat$", x)) + theme(plot.title = element_text(hjust = 0.5)))
}
l <- ls(xlabels)
```

```{r}
# Arrange and display the plots into a 2x4 grid. xlab(paste0("log( d$", x,")"))
grid.arrange(density_plot(l[1]),
             density_plot(l[2]),
             density_plot(l[3]),
             density_plot(l[4]), ncol=2)
```
```{r}
# Arrange and display the plots into a 2x4 grid
grid.arrange(density_plot(l[5]),
             density_plot(l[6]),
             density_plot(l[7]),
             density_plot(l[8]), ncol=2)
```
# Transform the Data
```{r, echo=TRUE, warning=FALSE}
# apply log transformation -- 7 variables
dat$alk.phos <- log(dat$alk.phos)
dat$bili <- log(dat$bili)
dat$chol <- log(dat$chol)
dat$copper <- log(dat$copper)
dat$protime <- log(dat$protime)
dat$ast <- log(dat$ast)
dat$trig <- log(dat$trig)
# change data labels -- 7 variables
d$trt <- factor(d$drug, 
                     levels = c("1", "2"), 
                     labels = c("D-penicillamine", "placebo"))
d$sex <- factor(d$sex, 
                     levels = c("0", "1"), 
                     labels = c("male", "female"))
d$ascites <- factor(d$ascites, 
                    levels = c("0", "1"), 
                    labels = c("no", "yes"))
d$hepato <- factor(d$hepato, 
                    levels = c("0", "1"), 
                    labels = c("no", "yes"))
d$spiders <- factor(d$spider, 
                    levels = c("0", "1"), 
                    labels = c("no", "yes"))
d$edema <- factor(d$edema, 
                    levels = c("0", "0.5", "1"), 
                    labels = c("no edema and no diuretic therapy for edema", 
                               "edema present without diuretics, or edema resolved by diuretics",
                               "edema despite diuretic therapy"))
d$stage <- factor(d$stage,
                      levels = c("1", "2", "3", "4"),
                      labels = c("one", "two", "three", "four"))
```

# One sample_nonparametric estimation
## Kaplan-Meyer estimator
```{r}
fit.KM <- survfit(Surv(time, status) ~ stage, data = dat)
summary(fit.KM)
```

```{r}
plot(fit.KM, mark.time = TRUE,
     main = "Kaplan-Meier estimator",
     ylab = "Survival probability",
     xlab = "time (years)" ,col = 1:4)
legend("bottomleft", lty = 1, col = 1:4, legend = names(fit.KM$strata))
```


### Median survival
```{r}
fit.KM
```
### Test the possibility to survive more than 10 years
```{r}
summary(fit.KM, time = 10)
#as shown below 40.7%
```

# Nonparametric comparison of two samples
## The logrank test
```{r}
fit.logrank <- survdiff(Surv(time, status) ~ stage, data = dat)
fit.logrank

```

# Cox model
The proportional hazards assumption is supported by a non-significant relationship between the _Schoenfeld residuals_ and time. The two must be independent.
## Testing the Proportional-Hazards Assumption with Schoenfeld residuals
```{r}
res.cox <- coxph(Surv(time, status) ~ trt + age + sex + ascites + hepato + 
                   spiders + edema + bili + chol + albumin + copper + alk.phos + 
                   ast + trig + protime + stage , data = dat)
test.ph <- cox.zph(res.cox)
test.ph
```
The output above shows that the test is not statistically significant for each of the covariates, and the global test is also not statistically significant.


The Cox Model assumes that continuous covariates have a nonlinear form. We can test the nonlinearity assumption by plotting the _Martingale residuals_ against the continuous covariates. 
## Testing the Proportional-Hazards Assumption with Martingale residuals
```{r}
ggcoxfunctional(Surv(time, status) ~  age + bili + chol + albumin + copper + 
                  alk.phos + ast + trig + protime, data = dat)
```
The assumption of non linearity is supported

## Fit the Cox model
```{r}
fit.cox <- coxph( Surv(time, status)  ~ sex + age + stage + trt + albumin + alk.phos + ascites + ast + bili+ chol+copper+edema+hepato+platelet+protime+sex+spiders+trig , data = dat)
summary(fit.cox)
```
```{r}
broom::tidy(fit.cox)
```
## Automatic model selection based on AIC
```{r}
Mfull <- coxph(Surv(time, status) ~ sex + age + stage + trt + albumin + alk.phos + ascites + ast + bili+ chol+copper+edema+hepato+platelet+protime+spiders+trig, data = train)

```

```{r}
MAIC <- step(Mfull)
```

```{r}
summary(MAIC)
# where we can see the 6 most important covariates by p-value
```
```{r}
MAIC[1]
# where we can see the 6 most important covariates by p-value
```
```{r}
#try 6 most important variables
try<-coxph(Surv(time, status) ~  age + stage + albumin +  bili+ chol+edema, data = train)
try
summary(try)

```


